---
title: "STATS 509 HOMEWORK 8"
author: "Yuan Yin"
output: pdf_document
font size: 11pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 2 (a)

### (i)

```{r message = FALSE, warning = FALSE}
library(Ecdat)
data(CRSPday)
crsp=CRSPday[,7]
ar1 = arima(crsp,order=c(1,0,0)); ar1
ar2 = arima(crsp,order=c(2,0,0)); ar2
```

+ I would prefer AR(1) model since we can see that the coefficient of $\phi_2$ in AR(2) is -0.014 and the standard error is 0.02, which means that $\phi_2 = 0$ is in the 95% confidence interval, this coefficient is not significant. Since AR(1) is a simpler model, so we prefer AR(1). What's more, AR(1) has smaller AIC, it corresponds with our result.

### (ii)

```{r}
left = ar1$coef[1]-1.96*sqrt(ar1$var.coef[1,1])
right = ar1$coef[1]+1.96*sqrt(ar1$var.coef[1,1])
c(left, right)
```

+ Seeing that the 95% interval of $\phi$ in AR(1) is $0.085\pm1.96*SE=[0.0465,0.1241]$.

### (iii)

```{r}
# Diagnostics plots from arima
tsdiag(ar1)
tsdiag(ar2)
```

```{r fig.height=4, fig.width=6}
## QQ plots of residuals
par(mfrow = c(1,2))
qqnorm(ar1$resid); qqline(ar1$resid)
qqnorm(ar2$resid); qqline(ar1$resid)
```

Seeing that both models fail to reject that residuals are white noise, and both QQ-plots show that residuals have heavy tails. This indicates that we can't tell differences between two models. For simplifying our models, we should choose AR(1). This doesn't change our answer

## Question 2 (b)

```{r message = FALSE, warning = FALSE}
library(forecast)
auto.arima(crsp, start.p=0, max.p = 4, max.q = 4, seasonal = FALSE, ic = "aic")
```

Seeing that MA(1) has the smallest AIC, let's make diagnostics on MA(1).

```{r}
ma1 = arima(crsp,order=c(0,0,1)); ma1
tsdiag(ma1)
qqnorm(ma1$resid); qqline(ma1$resid)
```

+ Finding that with similar results with AR(1) and AR(2), we still fail to reject the residuals to be white noise, but from QQ-plot we can see the distribution is not normal since its heavy tails.
+ Since MA(1) is the best AMRA model to fit the data, but there is not much difference between MA(1) and AR(1), maybe AMRA model is not a good model to fit the data, or maybe the residuals are nor normal distribution white noise but other type of white noise.