---
title: "STATS 531 Homework 6"
author: "Yuan Yin"
date: "3/15/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 7,1

First let's read in the data and see what they look like.

```{r message = FALSE, warning = FALSE}
dat <- read.csv("https://ionides.github.io/531w18/10/parus.csv")
require(pomp)
require(ggplot2)
set.seed(2018531)
parus = pomp(dat,times="year",t0=1959)
plot(parus)
```

Next, we need to add in the deterministic skeleton, so that R will know what our map is like (from $P_n$ to $P_{n+1}$).

```{r message = FALSE, warning = FALSE}
skel<-Csnippet("DN=r*N*exp(-N/k);")
parus<-pomp(parus,skeleton=map(skel),paramnames=c("r","k"),statenames=c("N"))
```

We plot the trajectory of our determinstic skeleton.

```{r}
traj<-trajectory(parus,params=c(N.0=1,r=12,k=1),as.data.frame=TRUE)
ggplot(data=traj,aes(x=time,y=N))+geom_line()
```

Now let's simulate data y with our pomp model.

```{r message = FALSE, warning = FALSE}
stochStep <- Csnippet("
  e = rnorm(0,sigma);
  N = r*N*exp(-N/k+e);
")
pomp(parus,rprocess=discrete.time.sim(step.fun=stochStep,delta.t=1),
     paramnames=c("r","k","sigma"),statenames=c("N","e")) -> parus
```

```{r}
sim <- simulate(parus,params=c(N.0=1,e.0=0,r=12,k=1,sigma=0.5),
                as.data.frame=TRUE,states=TRUE)
plot(N~time,data=sim,type='o')
lines(N~time,data=traj,type='l',col='red')
```

Seeing that there are still much difference between simulation of $P_n$ and original one. Now we use our model to get the measurement model of $Y_n$.

```{r}
rmeas <- Csnippet("pop = rnbinom(phi*N*psi/(1-psi),psi);")
dmeas <- Csnippet("lik = dnbinom(pop,phi*N*psi/(1-psi),psi,give_log);")
```

```{r message = FALSE, warning = FALSE}
pomp(parus,rmeasure=rmeas,dmeasure=dmeas,statenames=c("N"),paramnames=c("phi","psi")) -> parus
coef(parus) <- c(N.0=100,e.0=0,r=6,k=100,sigma=0.01,phi=1,psi=0.06)
```

```{r}
sims <- simulate(parus,nsim=3,as.data.frame=TRUE,include.data=TRUE)
ggplot(data=sims,mapping=aes(x=time,y=pop))+geom_line()+
  facet_wrap(~sim)
```

$\phi=1$ is used so that the mean of $Y_n$ equals to $P_n$. Moreover, the mean and variance of the dataset are used to determine $\psi$ (setting it to 0.06~mean(pop)/var(pop)). Set $r=6$, $k=100$, and $\sigma=0.01$. It seems like $r$ and $k$ determine the overall level of the pop.

## Question 8.2

With the same process above, we do all things again.

```{r}
skel<-Csnippet("DN=a*N/(1+b*N);")
parus<-pomp(parus,skeleton=map(skel),paramnames=c("a","b"),statenames=c("N"))
```

```{r}
traj<-trajectory(parus,params=c(N.0=1,a=4,b=2),as.data.frame=TRUE)
ggplot(data=traj,aes(x=time,y=N))+geom_line()
```

```{r}
stochStep <- Csnippet("
  e = rlnorm((-1)*sigma*sigma/2,sigma);
  N = a*N*e/(1+b*N);
")
pomp(parus,rprocess=discrete.time.sim(step.fun=stochStep,delta.t=1),
     paramnames=c("a","b","sigma"),statenames=c("N","e")) -> parus
```

```{r}
sim <- simulate(parus,params=c(N.0=1,e.0=1,a=4,b=2,sigma=0.2),
                as.data.frame=TRUE,states=TRUE)
plot(N~time,data=sim,type='o')
lines(N~time,data=traj,type='l',col='red')
```

```{r}
rmeas <- Csnippet("pop = rnbinom(phi*N*psi/(1-psi),psi);")
dmeas <- Csnippet("lik = dnbinom(pop,phi*N*psi/(1-psi),psi,give_log);")
```

```{r}
pomp(parus,rmeasure=rmeas,dmeasure=dmeas,statenames=c("N"),paramnames=c("phi","psi")) -> parus
coef(parus) <- c(N.0=100,e.0=1,a=400,b=2,sigma=0.2,phi=1,psi=0.06)
```

```{r}
sims <- simulate(parus,nsim=3,as.data.frame=TRUE,include.data=TRUE)
ggplot(data=sims,mapping=aes(x=time,y=pop))+geom_line()+
  facet_wrap(~sim)
```

Similarity with Ricker model: There is parameters $a$ and $b$ that could be used to determine the overall level of the population. With some choices of the parameters, the simulation results that have similar mean and variance with the dataset could be obtained.

Difference: When $\epsilon_n=1$ for all $n$, deterministic skeleton of $P_n$ does not show oscillatory behavior for any $a$ and $b$ for Beverton-Holt model. When $a$ and $b$ both positive, $P_n$ only grows (a plot showing this behavior is shown above).   

## Question 8.3

This homework takes about 2 days to finish.

+ Resource: Class notes and previous solution
+ Need to explain: The class notes and previous solution helps a lot, but I'm still confused about explaining the first plot about our simulation and original traj values. It seems like they are different from each other a lot, but why we still do it like this way.